---
title: Report of 2023 Research Activities 
author:
- affiliation: RACE
  description: Fish Biologist
  email: Alexandra.Dowlin@NOAA.gov
  name: Alexandra Dowlin
output: 
  word_document: 
    reference_docx: template-adfg.docx
fontsize:
addr:
  l1: 7600 Sand Pointe Way, NE
  l2: NMFS RACE Division, Groundfish Assessment Program
  l3: Seattle, WA 98115
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE)
```

State of Alaska Fish Resource Permit CF-22-022  

NOAA Fisheries Alaska Fisheries Science Center -- RACE Division  


# **Activity Title:**  
RACE Groundfish Assessment Program Biennial Bottom Trawl Survey of `r region` Groundfish and Invertebrate Resources  

# **Introduction:**  
The Alaska Fisheries Science Center (AFSC) Resource Assessment and Conservation Engineering (RACE) Groundfish Assessment Program's (GAP) bottom trawl survey of the `r regionw_abbr` is conducted to monitor the distribution, abundance, and biological characteristics of multiple species of groundfishes and invertebrates in coastal and offshore Alaska waters. The survey collects catch per unit effort (CPUE) of key species and uses the CPUE to estimate relative biomass from a stratified-random survey design. Survey data are also used to characterize length and age distributions, growth, recruitment, stock structure, and other important life history information of key species to better understand the ecological and physical factors affecting their distribution and abundance. These data, estimates, and study results are provided to federal and state stock assessment biologists and to the North Pacific Fishery Management Council (NPFMC) who use the information to determine productivity and to set annual Allowable Biological Catch (ABC) and Total Allowable Catch (TAC) for groundfishes in the `r region_abbr`.  

# **Methods:**  
The AFSC GAP biennial bottom trawl survey of the `r region_abbr` typically deploys two chartered commercial trawlers to conduct the survey for 75 days each, sampling the standard survey area (~320,000 km^2^) and trawling 520-550 survey stations. The `r region_abbr` survey uses a stratified-random design to allocate stations across the region in depths up to 700 m.  Survey strata are based on five depth intervals (1-100 m, 101-200 m, 201-300 m, 301-500 m and 501-700 m) and established survey districts and subdistricts. In `r maxyr`, the `r region_abbr` survey area is stratified into five International North Pacific Fisheries Commission (INPFC) statistical districts: Shumagins, Chirikof, Kodiak, Yakutat, and southeast Alaska. A grid of 5x5 km is superimposed on bathymetry maps with a grid cell representing the primary sampling unit. Stations determined to be untrawlable in previous surveys of the area are not considered when allocating survey stations to vessels. A Poly Nor’Eastern research net with roller gear is trawled on bottom at each station for up to 15 minutes. Trawl height, trawl width, and bottom contact are monitored with acoustic net mensuration gear; surface and bottom temperatures are collected using a bathythermograph attached to the headrope of the trawl as well. Catches are processed by identifying all animals collected to the highest possible taxonomic classification and weighing and counting each taxon. Additional samples are collected by measuring the length of selected species, collecting age structures, tissue samples, and stomach contents from selected specimens, saving voucher specimens for later identification and accessioning to museum collections. All data are preserved digitally and are subjected to multiple quality control checks.  

# **Results:**
Active fishing for the `r maxyr` `r region_abbr` bottom trawl survey occurred between `r dates_conducted` and was the `r survnumber`, standardized multi-species survey in the `r time_series` that begun in `r series_begun`. This also comprises the standardized stock assessment data set for the `r region_abbr`. Two charter vessels, the *`r vessel1`* and the *`r vessel2`*, were used to conduct the `r region_abbr` survey in summer of 2023 and the two vessels successfully trawled `r s_stations` stations out of `r t_stations` planned stations; `r u_hauls` of the `r a_hauls` trawls attempted were unsuccessful. A total of `r stations_3nm` trawls were attempted within Alaska state waters and collected `r fish_3nm` fish taxa weighing `r scales::comma(fish_3nm_wt)` kg (Table 1). From combined state and offshore waters, the survey collected `r fish_all` fish taxa weighing `r scales::comma(fish_all_wt)` kg in total (Table 2). There were `r inverts_3nm` invertebrate taxa collected from Alaska state waters on the `r maxyr` `r region_abbr` bottom trawl survey with a total weight of `r scales::comma(inverts_3nm_wt)` kg (Table 1). In total, `r inverts_all` invertebrate taxa were collected from combined state and offshore waters and weighed `r scales::comma(inverts_all_wt)` kg (Table 2). The same 6 fish species comprise 74% of the total catch weight from combined state and offshore waters and from within Alaska state waters alone in `r maxyr`. Those species are Pacific ocean perch, arrowtooth flounder, walleye pollock, Pacific halibut, Pacific cod, and flathead sole. In total, from all state and offshore waters sampled, the survey vouchered `r voucher_all_count` fish and invertebrate taxa and collected `r scales::comma(oto_all)` otoliths pairs from `r oto_taxa` fish species (Table 4). Of the `r voucher_all_count` vouchered taxa, 8 were fishes and the remaining 25 were invertebrates. Within Alaska state waters, `r voucher_3nm_count` invertebrates were vouchered and `r scales::comma(oto_3nm)` otolith pairs were collected from `r oto_3nm_taxa` fish species (Table 3). Finalized and validated survey data were released on `r data_finalized`. Data products (e.g., biomass estimates, length composition, and age composition) were presented to the NPFMC's Joint Plan Team during their annual September meeting. Survey data products have also been made available over the internet on the Alaska Fisheries Information Network (AKFIN^[https://akfin.psmfc.org/]) and the NMFS Fisheries One Stop Shop (FOSS^[https://www.fisheries.noaa.gov/foss/f?p=215:28]). 


\newpage
<!-- Catch Summary Table - Within 3-mile Territorial Waters -->
```{r, catch-summary-table}
library(here)
catch_summary_3nm1 = catch_summary_3nm 
#catch_summary_3nm1[is.na(catch_summary_3nm1)] <- ""
catch_sum_3nm <- catch_summary_3nm1 %>% arrange(desc(total_weight_kg)) %>%
  separate(species_name, c("species", "species_suffix"),
    sep = "(?=sp.)", extra = "merge", remove = FALSE
  ) %>%
  mutate(species_suffix = case_when(
    stringr::str_count(species, "\\w+") == 1 & is.na(species_suffix) ~ species,
    TRUE ~ species_suffix
  )) %>%
  separate(species, c("species_2", "egg_case"), sep = "(?=egg case)", remove = TRUE) %>%
  mutate(new_suffix = coalesce(species_suffix, egg_case)) %>%
  dplyr::select(-species_suffix, -egg_case)


for (i in 1:nrow(catch_sum_3nm)) {
  if (!is.na(catch_sum_3nm$new_suffix[i])) {
    if (catch_sum_3nm$species_2[i] == catch_sum_3nm$new_suffix[i]) {
      catch_sum_3nm$species_2[i] <- ""
    }
  }
}

flextable::flextable(catch_sum_3nm, col_keys = c("common_name", "dummy_col", "total_weight_kg", "total_count")) %>% 
  flextable::mk_par(j = "dummy_col",
    value = as_paragraph(
      as_i(species_2),
      "", new_suffix
  )) %>%
  flextable::add_header_lines("Table 1. 2023 Catch summary table - totals within 3-mile territorial waters (Alaska state waters)", top = TRUE) %>%
  flextable::theme_vanilla() %>%
  flextable::set_header_labels(
    common_name = "Common name",
    dummy_col = "Scientific name",
    total_weight_kg = "Weight in kg",
    total_count = "Number of individuals"
  )%>%
  flextable::width(width = 2.4, j = 1:2) %>%
  flextable::width(width = 1.1, j = 3:4) %>%
  flextable::align(align = "right", j = 3:4, part = "all") %>%
  flextable::align(align = "left", j = 1:2, part = "all") 
```


\newpage

<!-- **r maxyr** Catch Summary Table -- Totals for Survey (including state waters) -->

```{r, catch-summary-table2}
library(here)
catch_summary1 = catch_summary
#catch_summary1[is.na(catch_summary1)] <- ""
catch_sum_all <- catch_summary1 %>% 
  arrange(desc(total_weight_kg)) %>%
  separate(species_name, c("species", "species_suffix"),
    sep = "(?=sp.)", extra = "merge", remove = FALSE
  ) %>%
  mutate(species_suffix = case_when(
    stringr::str_count(species, "\\w+") == 1 & is.na(species_suffix) ~ species,
    TRUE ~ species_suffix
  )) %>%
  separate(species, c("species_2", "egg_case"), sep = "(?=egg case)", remove = TRUE) %>%
  mutate(new_suffix = coalesce(species_suffix, egg_case)) %>%
  dplyr::select(-species_suffix, -egg_case)


for (i in 1:nrow(catch_sum_all)) {
  if (!is.na(catch_sum_all$new_suffix[i])) {
    if (catch_sum_all$species_2[i] == catch_sum_all$new_suffix[i]) {
      catch_sum_all$species_2[i] <- ""
    }
  }
}


flextable::flextable(catch_sum_all, col_keys = c("common_name", "dummy_col", "total_weight_kg", "total_count")) %>% 
  flextable::mk_par(j = "dummy_col",
    value = as_paragraph(
      as_i(species_2),
      "", new_suffix
  )) %>% 
  flextable::add_header_lines("Table 2. 2023 Catch summary table - totals for survey (including state waters)", top = TRUE) %>%
  flextable::theme_vanilla() %>%
  flextable::set_header_labels(
    common_name = "Common name",
    dummy_col = "Scientific name",
    total_weight_kg = "Weight in kg",
    total_count = "Number of individuals"
  ) %>%
  flextable::width(width = 2.4, j = 1:2) %>%
  flextable::width(width = 1.1, j = 3:4) %>%
  flextable::align(align = "left", j = 1:2, part = "all") %>%
  flextable::align(align = "right", j = 3:4, part = "all")
```



\newpage

<!-- **r maxyr** Voucher or Collection Specimens from Alaska State Waters  -->

```{r, voucher-collection}
library(here)
voucher_3nm1 = voucher_3nm
voucher_3nm1[is.na(voucher_3nm1)] <- ""
voucher_3nm_srvy <- voucher_3nm1 %>%
  separate(species_name, c("species", "species_suffix"),
           sep = "(?=sp.)", extra = "merge", remove = FALSE
  ) %>%
  mutate(species_suffix = case_when(
    stringr::str_count(species, "\\w+") == 1 & is.na(species_suffix) ~ species,
    TRUE ~ species_suffix
  )) %>%
  separate(species, c("species_2", "egg_case"), sep = "(?=egg case)", remove = TRUE) %>%
  mutate(new_suffix = coalesce(species_suffix, egg_case)) %>%
  dplyr::select(-species_suffix, -egg_case)

for (i in 1:nrow(voucher_3nm_srvy)) {
  if (!is.na(voucher_3nm_srvy$new_suffix[i])) {
    if (voucher_3nm_srvy$species_2[i] == voucher_3nm_srvy$new_suffix[i]) {
      voucher_3nm_srvy$species_2[i] <- ""
    }
  }
}

flextable::flextable(voucher_3nm_srvy, col_keys = c("common_name", "dummy_col", "count", "comment")) %>% 
  flextable::mk_par(j = "dummy_col",
    value = as_paragraph(
      as_i(species_2),
      "", new_suffix
  ))  %>%
  flextable::add_header_lines("Table 3. 2023 Voucher or specimen collection within Alaska state waters", top = TRUE)  %>%
  flextable::theme_vanilla() %>%
  flextable::set_header_labels(
    common_name = "Common name",
    dummy_col = "Scientific name",
    count = "Record count",
    comment = "Collection type"
  ) %>%
  flextable::width(width = 1.1, j = 3:4) %>%
  flextable::width(width = 2.4, j = 1:2) %>%
  flextable::align(align = "right", j = 3:4, part = "all") %>%
  flextable::align(align = "left", j = 1:2, part = "all")
```


\newpage

<!-- **r maxyr** Voucher or Collection Specimens from All Waters --> 

```{r, voucher-all}
library(here)
voucher_all1 =  voucher_all
voucher_all1[is.na(voucher_all1)] <- ""
voucher_all_srvy <- voucher_all1 %>%
  #arrange(desc(count)) %>%
  separate(species_name, c("species", "species_suffix"),
           sep = "(?=sp.)", extra = "merge", remove = FALSE
  ) %>%
  mutate(species_suffix = case_when(
    stringr::str_count(species, "\\w+") == 1 & is.na(species_suffix) ~ species,
    TRUE ~ species_suffix
  )) %>%
  separate(species, c("species_2", "egg_case"), sep = "(?=egg case)", remove = TRUE) %>%
  mutate(new_suffix = coalesce(species_suffix, egg_case)) %>%
  dplyr::select(-species_suffix, -egg_case)

for (i in 1:nrow(voucher_all_srvy)) {
  if (!is.na(voucher_all_srvy$new_suffix[i])) {
    if (voucher_all_srvy$species_2[i] == voucher_all_srvy$new_suffix[i]) {
      voucher_all_srvy$species_2[i] <- ""
    }
  }
}

flextable::flextable(voucher_all_srvy, col_keys = c("common_name", "dummy_col", "count", "comment")) %>% 
  flextable::mk_par(j = "dummy_col",
    value = as_paragraph(
      as_i(species_2),
      "", new_suffix
  )) %>%
  flextable::add_header_lines("Table 4. 2023 Voucher or specimen collection for both state and offshore waters", top = TRUE) %>%
  flextable::theme_vanilla() %>%
  flextable::set_header_labels(
    common_name = "Common name",
    dummy_col = "Scientific name",
    count = "Record count",
    comment = "Collection type"
  ) %>%
  flextable::width(width = 1.1, j = 3:4) %>%
  flextable::width(width = 2.4, j = 1:2) %>%
  flextable::align(align = "right", j = 3:4, part = "all") %>%
  flextable::align(align = "left", j = 1:2, part = "all")
```